### 2016-01-21 04-evaluation.tex
Polish a bit on 02, 04, 04, 05

@@

\begin{figure}[b] % [htb]
  \centering
  \epsfig{file=plots/sweep-over-np-fixed-K.eps, width=0.48\textwidth}
  \caption{(a) Execution time of 2048 algorithm iterations for the
  same problem size (com-Friendster, K=1024, M=16384 n=32) across different
  cluster sizes. (b) Speedup achieved for same experiments in (a) with respect to
  8 nodes}
  \label{fig-strong-scaling}
\end{figure}

@@

\begin{figure}[htbp] % [htb]
  \centering
  \epsfig{file=plots/sweep-over-K-proportional-np.eps, width=0.48\textwidth}
  \caption{Average execution time per algorithm iteration varying
  the number of latent communities proportionally to the number of compute
  nodes.}
  \label{fig-weak-scaling}
\end{figure}

A study of a system's weak scaling aids in the assessment of the communication
and synchronization overheads associated with the management of large clusters.
Similarly, it can expose complex issues that hinder performance at scale such
as load imbalance. To fairly evaluate the algorithm's weak scaling behavior we
conducted several executions varying the cluster size and the number of latent
communities proportionally. The number of communities at 64 workers was 12K.
This methodology ensures that each compute node
performs a relatively constant amount of computational work across all
configurations. However, the number and size of messages exchanged between the
nodes would vary significantly. Figure~\ref{fig-weak-scaling} presents the
average execution
time per algorithm iteration across different cluster sizes.
This figure shows that the system's overall
overhead is minimal. Additionally, the experimental results show that the
implementation is capable of achieving good speedups provided the input problem
is large enough for the given cluster size.

@@

\begin{table}[htbp]
  \centering
  \begin{tabular}{l d{5.1} d{5.1}}
    \textbf{Iteration (sub)stages}  & \multicolumn{1}{c}{non-pipelined}
                                          & \multicolumn{1}{c}{pipelined} \\
    \hline
    draw/deploy mini-batch     &  45.6  &  26.2 \\
    load $\pi$                 & 205    & 209 \\
    update $\phi$              &  74    &  74 \\
    \hline
    total                      & 331    & 241 \\
    \hline
  \end{tabular}
  \caption{Stages that profit from the two pipelining optimizations.
  com-Friendster on 65
  compute nodes, with 12K communities. Times in ms per iteration.}
  \label{table-pipeline}
\end{table}

Table~\ref{table-pipeline} shows a breakdown of the (sub)stages that are
overlapped by both pipelining optimizations. The left-hand column has
all pipelining disabled, the right-hand
column has pipelining enabled. Within \textit{update\_phi},
loading $\pi$ from the DKV store is domininant.
The pipelining optimization significantly overlaps loading $\pi$ both with generation and
deployment of the mini-batch, and with the calculation of \textit{update\_phi}.

\begin{figure}[htb]
  \centering
  \epsfig{file=plots/sweep-over-K-fixed-np.eps, width=\columnwidth}
  \caption{Performance effect of varying the number of communities on the
    algorithm's execution time on 64 nodes when disabling or enabling both
    pipelining optimizations.}
  \label{fig-pipeline}
\end{figure}

@@

%################################################
\subsection{Horizontal vs. Vertical Scalability}
\begin{figure}[b] % [htb]
  \centering
  \epsfig{file=plots/hpc-cloud.eps, width=0.48\textwidth}
  \caption{Performance comparison between the distributed implementation
  running on DAS5 and the multi-threaded solution on machine with 40 cores and
  1TB of RAM.}
  \label{fig-scale-up}
\end{figure}
One of the main drawbacks of designing a distributed solution for a given
algorithm is the inherent complexity of communication and synchronization. A
considerably simpler approach would be developing a multi-threaded version and
running it on a machine with abundant memory and CPU cores. In such a context,
access to all of the algorithm's state would be an order of magnitude faster
than RDMA. Additionally, the overhead associated with synchronizing threads is
negligible compared to using MPI primitives. To evaluate the efficacy of both
approaches we utilized SURFsara's HPC Cloud system to instantiate a virtual
machine with 40 Intel Xeon E7-4850 cores clocked at 2.00GHz cores and 1TB of
memory. The physical machine underlying
the HPC Cloud system contains 40~CPU cores and does not oversubscribe
resources. Therefore, by provisioning all 40~cores we ensured that there is no
resource contention from other users of the system. Figure~\ref{fig-scale-up}
reports the execution time per algorithm iteration for two experimental setups.
%
First, Figure~\ref{fig-scale-up}-a shows the performance of executing the
algorithm on the HPC Cloud
system with 40 and 16 cores compared to a single DAS5 node with 16~cores. This
test uses the com-DBLP data set to enable the use of a large number of
communities without running out of memory. It is clear from the results that
the performance can benefit from the additional cores provided by HPC Cloud system.
%
Further, Figure~\ref{fig-scale-up}-b tests the performance of the HPC Cloud
system compared to 64 nodes of the DAS5
cluster using the com-Friendster data set. Clearly, the parallel and
distributed implementation vastly
outperforms the single-node multi-threaded solution. Moreover, the trajectory
of both curves shows a widening gap between them suggesting that the relative
performance difference will increase for larger $K$.
%
In conclusion, the overhead of network communication in the distributed version
is more than compensated by the increasing compute power compared to
a single-node implementation.

@@

\begin{figure}[b] % [htbp]
  \centering
  \epsfig{file=plots/qperf.eps, width=0.5\textwidth}
  \caption{Performance of DKV store compared to qperf.}
  \label{fig-qperf}
\end{figure}

>>>>

\begin{figure}[t] % [htbp]
  \centering
  \epsfig{file=plots/sweep-over-K-proportional-np.eps, width=0.48\textwidth}
  \caption{Average execution time per algorithm iteration when varying
  the number of latent communities proportionally to the number of compute
  nodes.}
  \label{fig-weak-scaling}
\end{figure}

A study of a system's weak scaling aids in the assessment of the communication
and synchronization overheads associated with the management of large clusters.
Similarly, it can expose complex issues that hinder performance at scale such
as load imbalance. To fairly evaluate the algorithm's weak scaling behavior we
conducted several executions varying the cluster size and the number of latent
communities proportionally. The number of communities is 192 times the number
of worker nodes in all runs.
This methodology ensures that each compute node
performs an approximately constant amount of computational work across all
configurations. However, the number and size of messages exchanged between the
nodes would vary significantly. Figure~\ref{fig-weak-scaling} presents the
average execution
time per algorithm iteration across different cluster sizes.
This figure shows that the system's overall
overhead is minimal. Additionally, the experimental results show that the
implementation is capable of achieving good speedups provided the input problem
is large enough for the given cluster size.

@@

\begin{figure}[t] % [htb]
  \centering
  \epsfig{file=plots/sweep-over-np-fixed-K.eps, width=0.48\textwidth}
  \caption{(a) Execution time of 2048 algorithm iterations for the
  same problem size (com-Friendster, K=1024, M=16384 n=32) across different
  cluster sizes. (b) Speedup achieved for same experiments in (a) with respect to
  8 nodes}
  \label{fig-strong-scaling}
\end{figure}

@@

@@

\begin{table}[t] % [htbp]
  \centering
  \begin{tabular}{l d{5.1} d{5.1}}
    \textbf{Iteration (sub)stages}  & \multicolumn{1}{c}{non-pipelined}
                                          & \multicolumn{1}{c}{pipelined} \\
    \hline
    draw/deploy mini-batch     &  45.6  &  26.2 \\
    load $\pi$                 & 205    & 209 \\
    update $\phi$              &  74    &  74 \\
    \hline
    total                      & 331    & 241 \\
    \hline
  \end{tabular}
  \caption{Stages that profit from the two pipelining optimizations.
  com-Friendster on 65
  compute nodes, with 12K communities. Times in ms per iteration.}
  \label{table-pipeline}
\end{table}

Table~\ref{table-pipeline} shows a breakdown of the (sub)stages that are
overlapped by both pipelining optimizations. The left-hand column has
all pipelining disabled, the right-hand
column has pipelining enabled. Within \textit{update\_phi},
loading $\pi$ from the DKV store is dominant.
The pipelining optimization significantly overlaps loading $\pi$ both with generation and
deployment of the mini-batch, and with the calculation of \textit{update\_phi}.

\begin{figure}[t] % [htb]
  \centering
  \epsfig{file=plots/sweep-over-K-fixed-np.eps, width=\columnwidth}
  \caption{Performance effect of varying the number of communities on the
    algorithm's execution time on 64 nodes when disabling or enabling both
    pipelining optimizations.}
  \label{fig-pipeline}
\end{figure}

@@

%################################################
\subsection{Horizontal vs. Vertical Scalability}
\begin{figure}[t] % [htb]
  \centering
  \epsfig{file=plots/hpc-cloud.eps, width=0.48\textwidth}
  \caption{Performance comparison between the distributed implementation
  running on DAS5 and the multi-threaded solution on machine with 40 cores and
  1TB of RAM.}
  \label{fig-scale-up}
\end{figure}
One of the main drawbacks of designing a distributed solution for a given
algorithm is the inherent complexity of communication and synchronization. A
considerably simpler approach would be developing a multi-threaded version and
running it on a machine with abundant memory and CPU cores. In such a context,
access to all of the algorithm's state would be an order of magnitude faster
than RDMA. Additionally, the overhead associated with synchronizing threads is
negligible compared to using MPI primitives. To evaluate the efficacy of both
approaches we utilized SURFsara's HPC Cloud system to instantiate a virtual
machine with 40 Intel Xeon E7-4850 cores clocked at 2.00GHz cores and 1TB of
memory. The physical machine underlying
the HPC Cloud system contains 40~CPU cores and does not oversubscribe
resources. Therefore, by provisioning all 40~cores we ensured that there is no
resource contention from other users of the system. Figure~\ref{fig-scale-up}
reports the execution time per algorithm iteration for two experimental setups.
%
First, Figure~\ref{fig-scale-up}-a shows the performance of executing the
algorithm on the HPC Cloud
system with 40 and 16 cores compared to a single DAS5 node with 16~cores. This
test uses the com-DBLP data set to enable the use of a large number of
communities without running out of memory. It is clear from the results that
the performance can benefit from the additional cores provided by HPC Cloud system.
%
Further, Figure~\ref{fig-scale-up}-b tests the performance of the HPC Cloud
system compared to 64 nodes of the DAS5
cluster using the com-Friendster data set. Clearly, the parallel and
distributed implementation vastly
outperforms the single-node multi-threaded solution. Moreover, the trajectory
of both curves shows a widening gap between them suggesting that the relative
performance difference will increase for larger $K$.
%
In conclusion, the overhead of network communication in the distributed version
is more than compensated by the increasing compute power compared to
a single-node implementation.

@@

\begin{figure}[t] % [htbp]
  \centering
  \epsfig{file=plots/qperf.eps, width=0.5\textwidth}
  \caption{Performance of DKV store compared to qperf.}
  \label{fig-qperf}
\end{figure}

### 2016-01-21 05-conclusion.tex
More little textual thingies
Polish a bit on 02, 04, 04, 05

We conducted a thorough empirical evaluation of the system to study its strong
and weak scalability on 64 cluster nodes using large data sets.  Additionally,
we assessed the efficiency of the algorithm's resource utilization. Finally, a
demonstration of the implementation's utility was provided by processing 6
different organic data sets.

>>>>

We conducted a thorough empirical evaluation of the system to study its strong
and weak scalability on 65 cluster nodes using large data sets.  Additionally,
we assessed the efficiency of the algorithm's resource utilization. Finally, a
demonstration of the implementation's utility was provided by processing 6
different organic data sets.
To the best of
our knowledge, this is the first time that the problem of deducing overlapping
communities has been learned for problems of such a large scale.
